\begin{abstract}
EEG foundation models such as LaBraM~\cite{labram2024}, EEGPT~\cite{eegpt2024},
NeuroLM~\cite{neurolm2025}, and the concurrent work Laya~\cite{laya2026} pretrain
on tens of thousands of hours of EEG with 5\textendash 10\,M$+$ parameters and
multi-GPU training budgets, targeting cross-dataset generalization. The
complementary regime of \emph{focused single-corpus pretraining for in-
distribution downstream tasks}---realistic for individual research labs and
on-device deployment---remains underexplored. We present
\textbf{EEG-LeJEPA}, a 2.86\,M-parameter Joint-Embedding Predictive
Architecture for EEG, trained with Sketched Isotropic Gaussian
Regularization (SIGReg)~\cite{lejepa2025}. Pretrained for 10{,}000 steps on
50 EEGMMIDB subjects (subjects~51--100, disjoint from evaluation) in 10
minutes on a single consumer GPU ($\sim$16$\times$ less compute than
concurrent work~\cite{laya2026}), the model achieves 68.6\,\% accuracy /
0.751 AUC on cross-subject rest-versus-activity classification and 62.1\,\% /
0.690 AUC on left-versus-right motor imagery via a frozen-encoder linear
probe on encoder mean-pooled features (LOSO with subjects~1--20 strictly
held out of pretraining). Both gains over a random-initialized identical
architecture are large and statistically significant under a paired
permutation test ($+11.7$ and $+10.9$\,pp; Benjamini--Hochberg
$q<0.002$). We characterize the scaling behaviour of this compact regime
across six (data, compute) operating points: best-source AUC rises and then
plateaus near 0.76. A 2.4$\times$ larger model is modestly but consistently
better at matched data ($+1.3$ to $+2.0$\,pp on encoder mean-pooled features,
seed-stable across three seeds), so the compact model is a deliberate
accuracy/size trade-off for edge deployment rather than a capacity-optimal
point. A SIGReg-weight ablation reveals a sharp
U-shaped optimum at $\lambda=1.0$ with complete representational collapse at
$\lambda=0$ (probe accuracy regresses to exactly chance). A counterintuitive
channel-transfer experiment shows that pretraining on the channel
intersection with the downstream dataset \emph{underperforms} channel-padded
transfer from the full-channel source. Our SSL$+$linear-probe pipeline
matches supervised-from-scratch training of the same architecture on
left-versus-right MI ($-0.2$\,pp, n.s.) and significantly exceeds it on
rest-versus-activity ($+8.4$\,pp, $p<0.001$), with one frozen encoder serving
both tasks. The pipeline is reproducible on a single consumer GPU in under
15 minutes.
\end{abstract}
